---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data  
### 1. Unzip the file  
```{r}
unzip("activity.zip", overwrite=TRUE)
```

### 2. Load the data  
```{r}
activity <- read.csv("activity.csv")
```

### 3. Read the summary of data  
```{r}
summary(activity)
```

### 4. Read the first 6 rows  
```{r}
head(activity)
```

## What is mean total number of steps taken per day?  
### 1. Calculate the total number of steps taken per day  
```{r}
library(plyr)
activity_total_steps <- ddply(activity, .(date), summarize, total_steps_pday = sum(steps, na.rm=TRUE))
head(activity_total_steps)
```

### 2. Make a histogram of the total number of steps taken each day  
```{r}
with(activity_total_steps, hist(total_steps_pday, xlab="Total number of steps per day", breaks=8, col="red"))
```

### 3. Calculate and report the mean and median of the total number of steps taken per day  
#### 3.1 Mean of the total number of steps taken per day  
```{r}
with(activity_total_steps, round(mean(total_steps_pday), digits = 0))
```

#### 3.2 Median of the total number of steps taken per day  
```{r}
with(activity_total_steps, median(total_steps_pday))
```

## What is the average daily activity pattern?  
### 1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)  
```{r}
daily_activity_pattern <- ddply(activity, .(interval), summarize, total_steps_pday = mean(steps, na.rm=TRUE))
with(daily_activity_pattern, plot(interval, total_steps_pday, type="l",  
                                  xlab="5-minute interval", 
                                  ylab="Average number of steps taken across all days"))
```

### 2. 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps  
```{r}
max_ind <- which.max(daily_activity_pattern$total_steps_pday)
daily_activity_pattern[max_ind, "interval"]
```

## Imputing missing values  
### 1. Calculate and report the total number of missing values in the dataset  
```{r}
nrow(subset(activity, is.na(steps)))
```

### 2. Filling in all of the missing values in the dataset using **the mean for that 5-minute interval**, create a new dataset that is equal to the original dataset but with the missing data filled in  
```{r}
activity_imputed <- join(activity, daily_activity_pattern, by="interval")
activity_imputed[is.na(activity_imputed$steps), "steps"] <- activity_imputed[is.na(activity_imputed$steps), "total_steps_pday"]
activity_imputed$total_steps_pday <- NULL
head(activity_imputed)
```

### 3. Make a histogram of the total number of steps taken each day  
```{r}
activity_total_steps <- ddply(activity_imputed, .(date), summarize, total_steps_pday = sum(steps, na.rm=TRUE))
with(activity_total_steps, hist(total_steps_pday, xlab="Total number of steps per day", breaks=8, col="red"))
```

### 4. Calculate and report the **mean** and **median** total number of steps taken per day  
#### 4.1 **Mean**  total number of steps taken per day  
```{r}
with(activity_total_steps, round(mean(total_steps_pday), digits = 0))
```

#### 4.2 **Median** total number of steps taken per day  
```{r}
with(activity_total_steps, round(median(total_steps_pday), digits = 0))
```

## Are there differences in activity patterns between weekdays and weekends?  
#### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day  
```{r}
activity_imputed$date <- as.Date(activity_imputed$date, format = "%Y-%m-%d")
activity_imputed$weekday <- as.factor(ifelse(weekdays(activity_imputed$date) %in% c("Saturday", "Sunday"), "weekend", "weekday"))
head(activity_imputed)
```

### 2. Make a panel plot containing a time series plotof the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)  
```{r}
# Calculate the average number of steps taken across all weekday days or weekend days
activity_imputed_activity_pattern <- ddply(activity_imputed, .(interval, weekday), summarize, total_steps_pday = mean(steps, na.rm=TRUE))
# load lattice library
library(lattice)
xyplot(total_steps_pday ~ interval | weekday, data= activity_imputed_activity_pattern, layout = c(1, 2), ylab = "Number of steps", 
       panel = function(x, y, ...) {
         panel.xyplot(x, y, type="l", ...)
         })